---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo = TRUE}
# Importing required libraries
library(dplyr)
library(ggplot2)
library(tidyr)

```




## Loading and preprocessing the data
```{r, echo = TRUE}
activity = read.csv(unz("activity.zip", "activity.csv"), stringsAsFactors = FALSE, na.strings = "NA")

```


## What is mean total number of steps taken per day?
```{r, echo = TRUE}

# Calculate the steps by date
steps_per_day = activity %>%
  group_by(date) %>%
  summarise(total_steps=sum(steps, na.rm=TRUE))

```


```{r, echo = TRUE}
# Generate histogram of the daily steps
ggplot(data=steps_per_day, aes(x=total_steps)) + geom_histogram() + labs(x='Total steps', y='Count')

```


```{r, echo = TRUE}
# Summarizing mean and median
steps_per_day %>% 
  summarise(
    daily_steps_mean=mean(total_steps),
    daily_steps_median=median(total_steps)
  )

```


## What is the average daily activity pattern?
```{r, echo = TRUE}
interval_mean = activity %>%
  group_by(interval) %>%
  summarize(mean_steps=mean(steps, na.rm=TRUE))


# Viewing the intervall mean over days
ggplot(data=interval_mean, aes(x=interval, y=mean_steps)) + geom_line() + labs(x='Interval', y='Mean steps')

```
```{r, echo = TRUE}

# Finding the interval with the greatest mean
interval_mean %>% 
  arrange(desc(mean_steps)) %>%
  slice(1)

```

The 835 interval was the interval were the steps were the maximum in this dataset.


## Imputing missing values

```{r, echo = TRUE}

# Counting na by column
sapply(activity, function(y) sum(length(which(is.na(y)))))

```

```{r, echo = TRUE}




new_activity = merge(activity, interval_mean, by='interval', all.x=TRUE)

# Fill na with mean steps in interval
new_activity$steps[is.na(new_activity$steps)] = new_activity$mean_steps

new_activity = new_activity %>%
  arrange(date, interval) %>%
  select(date, interval, steps)

sapply(new_activity, function(y) sum(length(which(is.na(y)))))

```

```{r, echo = TRUE}

# Calculate the steps by date
new_steps_per_day = new_activity %>%
  group_by(date) %>%
  summarise(total_steps=sum(steps))


# Generate histogram of the daily steps
ggplot(data=new_steps_per_day, aes(x=total_steps)) + geom_histogram() + labs(x='Total steps', y='Count')

```


```{r, echo = TRUE}
# Summarizing mean and median
new_steps_per_day %>% 
  summarise(
    daily_steps_mean=mean(total_steps),
    daily_steps_median=median(total_steps)
  )

```

For the mean the difference is low and for the median there is no change. Maybe it could be that the missing values were clustered where the steps count were low.


## Are there differences in activity patterns between weekdays and weekends?



```{r, echo = TRUE}

# Adding weekday as factor
new_activity= new_activity %>%
  mutate(date=as.Date(date)) %>%
  mutate(weekday=ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"), weekdays(date)) %>%
  mutate(weekday=as.factor(weekday))

# Generating aggregated data by weekday
steps_by_interval_weekday = new_activity %>%
  group_by(interval, weekday) %>%
  summarise(avg_steps=mean(steps, na.rm=TRUE))

# Plotting aggregated data
ggplot(data=steps_by_interval_weekday, aes(x=interval, y=avg_steps)) + labs(x='Interval', y='Number of steps') + geom_line() + facet_grid(rows=vars(weekday))

```


It looks that on the weekend the steps were more evenly distributed during the day, and on weekdays it peaks between 750-1000 and the remain in more lower values.
